---
title: "p8105_hw2_rl3328"
author: "Ruixi Li"
date: "2023-09-28"
output: github_document
---

# import and clean data
```{r setup, message=FALSE}
library(tidyverse)
library(readr)
library(readxl)
```


# Problem 1

## import and clean pols-month
```{r pols, message=FALSE}
pols = 
  read_csv("./data/fivethirtyeight_datasets/fivethirtyeight_datasets/pols-month.csv") |>
  janitor::clean_names() |>#this step can be ignored cuz the variable names are clean
  separate(
    col = mon, 
    into = c("year","month","day"), 
    sep = "-") |>#I used `remove = FALSE` to check if I have done it correctly(not shown)
  mutate(month = tolower(month.abb[as.numeric(month)]))|>#I assume month name is abbreviation
  mutate(president =recode(prez_gop, "0" = "dem", "1" = "gop", "2" = "gop")) |>
  select(-c("prez_gop","prez_dem","day"))




```

## import and clean snp

After knowing the range of years, I assign the date into correct century.
```{r snp, message=FALSE}
snp = 
  read_csv("./data/fivethirtyeight_datasets/fivethirtyeight_datasets/snp.csv") |>
  janitor::clean_names() |>#this step can be ignored cuz the variable names are clean
  separate(
    col = date, 
    into = c("month","day","year"), 
    sep = "/") |>#I used `remove = FALSE` to check if I have done it correctly(not shown)
  mutate(year = ifelse(as.numeric(year)>15, 1900+as.numeric(year), 2000+as.numeric(year)),
         month = tolower(month.abb[as.numeric(month)])) |>
  select(-"day")
```


## import and clean unemployment

```{r unemployment, message=FALSE}
unemployment = 
  read_csv("./data/fivethirtyeight_datasets/fivethirtyeight_datasets/unemployment.csv") |>
  janitor::clean_names() |>
  pivot_longer(
    -year,
    names_to = "month", 
    values_to = "unemployment")  
```

## merge three datasets together
Join the datasets by merging `snp`into `pols`, and merging `unemployment` into the result.

```{r merge}
snp = snp |> mutate(year = as.character(year))

results = left_join(pols, snp, by = c("year","month"))

unemployment = unemployment |> mutate(year = as.character(year))

results = left_join(results, unemployment, by = c("year","month"))

```


# Problem 2

## import and clean mr. transh wheel
```{r mr,message=FALSE}
mr = read_excel("./data/trash.xlsx", 1) |>
  janitor::clean_names() |>
  select(-c(15,16)) |>
  mutate(month = ifelse(month == "Decemeber","December",month),
         sports_balls = round(sports_balls))

str(mr)

summary(mr)

```
* there's one `NA` in `dumpster` variable, one `NA` in `date` variable and  61 `NAs` in `homes_powered` variable.
* I don't like the input file name with space, so I changed the name of it.

* delete this row whose dumpster is `NA`

```{r drop}
mr = drop_na(mr, dumpster)
```

* calculate homes powered

* according to `home powered note`, Homes Powered - Each ton of trash equates to on average 500 kilowatts of electricity.  An average household will use 30 kilowatts per day.


```{r}
mr = mr |> mutate(homes_powered = (weight_tons*500)/30,
                  trash_wheel_name = "Mr.Trash Wheel")
```
I can integrate the seperated codes into one using pipeline, but I am showing my process of thinking.

## import and clean Professor Trash Wheel and Gwynnda

```{r}
pro = read_excel("./data/trash.xlsx", 2) |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(homes_powered = (weight_tons*500)/30,
         trash_wheel_name = "Professor Trash Wheel",
         year = as.character(year))
  

str(pro)

summary(pro)



gwy = read_excel("./data/trash.xlsx", 4) |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(homes_powered = (weight_tons*500)/30,
         trash_wheel_name = "Gwynnda Trash Wheel",
         year = as.character(year),
         month = ifelse(month == "july","July",month))


str(gwy)
summary(gwy)
```


## combine them together

```{r}
trash =  bind_rows(mr,gwy,pro)
```

Let's talk about the Mr. Trash Wheel datasets. The `mr` data has `r nrow(mr)` observations and `r ncol(mr)` variables and tells us about the amount of trash the dumpster receives(measured by both weight `weight_tons` and volume `volume_cubic_yards`), fill time`month`,`year`,`date` and the categories of trash`r colnames(mr[,7:15])` for "Mr. Trash Wheel" from years `r range(mr$year)[1]` to `r range(mr$year)[2]`. It also tells us how many homes can be powered by the electricity produced by trash incineration. The `pro` data has `r nrow(pro)` observations and `r ncol(pro)` variables, ranging from years `r range(pro$year)[1]` to `r range(pro$year)[2]`. The `gwy` data has `r nrow(gwy)` observations and `r ncol(gwy)` variables ranging from years `r range(gwy$year)[1]` to `r range(gwy$year)[2]`.**The total weight of trash collected by Professor Trash Wheel was `r filter(trash, trash_wheel_name == "Professor Trash Wheel") |> pull(weight_tons) |> sum() |> round(2)`**. **The total number of cigarette butts collected by Gwynnda in July of 2021 was `r filter(trash, year == "2021", month == "July",trash_wheel_name == "Gwynnda Trash Wheel") |> pull(cigarette_butts) |> sum()`**.



# Problem 3

## import and clean baseline

```{r baseline_check, include=FALSE}
baseline = read_csv("./data/data_mci/MCI_baseline.csv",col_names = FALSE) |>
  slice(-c(1,2)) |>
  magrittr::set_colnames(c("ID","current_age","sex","education","apoe4","age_at_onset")) |>
  janitor::clean_names() |>
  mutate(
    apoe4_labled = recode(apoe4, "1" = "APOE4 carrier","0" = "APOE4 non-carrier"),
    sex_labled = recode(sex, "1" = "Male", "0" = "Female"))


table(baseline$apoe4,baseline$apoe4_labled)
table(baseline$sex,baseline$sex_labled)

```

```{r baseline,message=FALSE}
baseline = read_csv("./data/data_mci/MCI_baseline.csv",col_names = FALSE) |>
  slice(-c(1,2)) |>
  magrittr::set_colnames(c("ID","current_age","sex","education","apoe4","age_at_onset")) |>
  janitor::clean_names() |>
  mutate(
    apoe4 = recode(apoe4, "1" = "APOE4 carrier","0" = "APOE4 non-carrier"),
    sex = recode(sex, "1" = "Male", "0" = "Female"),
    current_age = as.numeric(current_age),
    age_at_onset = ifelse(age_at_onset == ".", NA, age_at_onset),
    age_at_onset = as.numeric(age_at_onset)) |>
  filter(current_age<age_at_onset | is.na(age_at_onset))

str(baseline)
summary(baseline)


prop.table(table(baseline$apoe4))

```
* When importing `baseline` data, I first drop the first row of the dataset(after memorizing the decoding rules) and then used the second row as column names. After cleaning the column names, I recoded `sex` and `apoe4`. Since `current_age` and `age_at_onset` are all character type in the original file. I converted them as numeric before comparing them to identify not MCI-free subjects at baseline and replace "." with `na`in age_at_onset. No duplicate `id` in `baseline` data.
* The `baseline` data has `r nrow(baseline)` observations and `r ncol(baseline)` variables and tells us about the demographic features at baseline(sex, current_age,apoe4 status, education and MCI status during follow-up). 
* `r nrow(baseline)` participants were recruited, and of these `r baseline |> filter(!is.na(age_at_onset)) |> nrow()` develop MCI. The average baseline age is `r baseline |> pull(current_age) |> mean()`. 30.1% of women in the study are APOE4 carriers.

## import and clean amyloid data

```{r amyloid, message=FALSE}
amyloid = read_csv("./data/data_mci/mci_amyloid.csv",col_names = FALSE) |>
  slice(-c(1,2)) |>
  magrittr::set_colnames(c("ID","abeta_40/42_bsl","abeta_40/42_2","abeta_40/42_4","abeta_40/42_6","abeta_40/42_8")) |>
  janitor::clean_names() |>
  mutate_at(c(2:6), as.numeric)
  
  
str(amyloid)
summary(amyloid)
```
* When importing `baseline` data, I first drop the first row of the dataset(after memorizing the decoding rules) and then used the second row(modified a little) as column names. I found there's "Na" in follow-up columns, so I converted them into numeric so that some strings unable to be converted directly became `NA` . No duplicate `id` in the `amyloid` data.

## check and combine

```{r}
a <- left_join(baseline,amyloid,by="id")
b <- right_join(baseline,amyloid,by="id")
merge <- inner_join(baseline,amyloid,by="id")
d <- full_join(baseline,amyloid,by="id")
```
* there's `r nrow(merge)` participants have both `baseline` and `amyloid` data; `r nrow(baseline)-nrow(merge)` participants only appear in `baseline` data; `r nrow(amyloid)-nrow(merge)` participants only appear in `amyloid` data.
* `merge` dataset has `r nrow(merge)` observations and `r ncol(merge)` variables and gave us `r nrow(merge)` participants' demographic features and their Amyloid 40/42 ratio during follow-up time point.

```{r}
write_csv(merge,"merge.csv")
```

